---
title: "Normality Testing in R"
author: "Dr. Lindokuhle Dlamini"
date: "12 March 2025"
output:
  html_document:
    toc: false
    include:
      title: false
      author: false
      date: false
    css: "styles.css"  # Ensure this path matches where your CSS is stored
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<header>
    <h1 class="title">Normality Testing in R</h1>
    <p>Visual and Statistical Approaches for Checking Normality</p>
    <h4 class="author"><em>Dr Lindokuhle Dlamini</em></h4>
    <h4 class="date"><em>12 March 2025</em></h4>
</header>
## Introduction
This document outlines the process for performing normality testing in R using visual and statistical methods. The dataset used here is based on plant richness.

## 1. Loading the Dataset
```{r}
getwd()
setwd("G:/My Drive/Rhodes/2025/Honours 2025/Intro to R Hons 2025")
library(readxl)
library(tidyverse)
library(ggplot2)
Independent_t_test <- read_excel("Independent t test.xlsx")

data2025 <- Independent_t_test  # Renaming the dataset for ease of use
View(data2025)
```

## 2. Visual Normality Testing
### 2.1 Histogram
```{r}
# Histogram
hist(data2025$`Plant richness`)
```

### 2.2 Enhanced Histogram with Normal Distribution Curve
```{r}
# Enhanced Histogram with normal curve
hist(data2025$`Plant richness`, 
     probability = TRUE,  
     col = "lightblue",   
     main = "Histogram with Normal Distribution", 
     xlab = "Plant Richness", 
     border = "black", 
     ylim = c(0, 0.100))

curve(dnorm(x, mean = mean(data2025$`Plant richness`), 
            sd = sd(data2025$`Plant richness`)), 
      col = "red", lwd = 3, add = TRUE)
```

### 2.3 Density Plot using ggplot2
```{r}
# Density plot using ggplot2
ggplot(data = data2025, aes(x = `Plant richness`)) + 
  geom_density() + theme_classic()
```

### 2.4 Density Plot using Base R
```{r}
# Density plot in base R
density_plot <- density(data2025$`Plant richness`)
plot(density_plot, 
     main = "Density Plot of Plant Richness", 
     xlab = "Plant Richness", 
     ylab = "Density", 
     col = "blue", 
     lwd = 2)
```

### 2.5 Q-Q Plot
```{r}
# Q-Q plot
qqnorm(data2025$`Plant richness`)
qqline(data2025$`Plant richness`, col = "red")
```

## 3. Statistical Normality Testing
### 3.1 Shapiro-Wilk Test
```{r}
# Shapiro-Wilk Test
shapiro.test(data2025$`Plant richness`)
```

### 3.2 Kolmogorov-Smirnov Test
```{r}
# Kolmogorov-Smirnov Test
ks.test(data2025$`Plant richness`, "pnorm", mean = mean(data2025$`Plant richness`), sd = sd(data2025$`Plant richness`))
```

## Footer
<style>
    body {
        font-family: Arial, sans-serif;
        line-height: 1.6;
        background-color: #f4f4f4;
        margin: 0;
        padding: 0;
    }
    header {
        background-color: #628554;
        color: #fff;
        text-align: center;
        padding: 10px;
    }
    h1, h2, h3, h4, h5, h6 {
        color: #333;
    }
    code {
        background-color: #d3f5e0;
        padding: 5px;
        border-radius: 3px;
        font-family: monospace;
    }
    pre {
        background-color: #f0f0f0;
        padding: 10px;
        border-radius: 5px;
        overflow-x: auto;
    }
    footer {
        text-align: center;
        padding: 10px;
        background-color: #d5e8cc;
        color: #fff;
        position: fixed;
        width: 100%;
        bottom: 0;
    }
</style>

<footer>
    <p>&copy; 2025 Dr. Lindokuhle Dlamini</p>
</footer>
