---
title: "Normality Test in R"
author: "Dr Lindokuhle Dlamini"
date: "2025-03-12"
output: html_document
---

```{r setup, include=FALSE}
getwd()
setwd("G:/My Drive/Rhodes/2025/Honours 2025/Intro to R Hons 2025")
library(readxl)
library(tidyverse)
library(ggplot2)
Independent_t_test <- read_excel("Independent t test.xlsx")
data2025 <- Independent_t_test  # Renaming the dataset for ease of use 
View(data2025)
```

Normality Testing Using Visuals
1. Histogram
You can use a histogram to visually inspect the distribution of data.

```{r}
# Basic Histogram
hist(data2025$`Plant richness`)

# Enhanced Histogram with Normal Distribution Curve
hist(data2025$`Plant richness`, 
     probability = TRUE,  # To scale the histogram to show probabilities
     col = "lightblue",   # Color of the histogram bars
     main = "Histogram with Normal Distribution", 
     xlab = "Plant Richness", 
     border = "black", 
     ylim = c(0, 0.100))   # Adjust the y-axis to ensure the curve fits

# Add a red line for the normal distribution curve
curve(dnorm(x, mean = mean(data2025$`Plant richness`), 
            sd = sd(data2025$`Plant richness`)), 
      col = "red", 
      lwd = 3, 
      add = TRUE)  # Overlay the curve on the histogram

```
2. Density Plot
We can create a density plot using both base R and ggplot2.

2.1 Density Plot with ggplot2
```{r}
# Density plot using ggplot2
ggplot(data = data2025, aes(x = `Plant richness`)) + 
  geom_density() + 
  theme_classic()
```
2.2 Density Plot with Base R

```{r}
# Create and plot density object in base R
density_plot <- density(data2025$`Plant richness`)
plot(density_plot, 
     main = "Density Plot of Plant Richness", 
     xlab = "Plant Richness", 
     ylab = "Density", 
     col = "blue",   # Color of the density curve
     lwd = 2)        # Line width

```

3. Q-Q Plot
A Q-Q plot is a more formal method for inspecting the normality of a dataset.

```{r}
# Q-Q plot for visual inspection of normality
qqnorm(data2025$`Plant richness`)
qqline(data2025$`Plant richness`, col = "red")

```
Normality Test Statistically
4. Shapiro-Wilk Test
The Shapiro-Wilk test is one of the most popular tests for normality. It tests the null hypothesis that the data is normally distributed.

```{r}
# Shapiro-Wilk Test
shapiro.test(data2025$`Plant richness`)

```
5. Kolmogorov-Smirnov Test
The Kolmogorov-Smirnov test compares the sample distribution to a reference distribution (such as the normal distribution).
```{r}
# Kolmogorov-Smirnov Test
ks.test(data2025$`Plant richness`, "pnorm", mean = mean(data2025$`Plant richness`), sd = sd(data2025$`Plant richness`))


```
Conclusion
In this analysis, we demonstrated how to visually and statistically test for normality in a dataset using R. First, we used histograms, density plots, and Q-Q plots for visual inspection. Then, we performed the Shapiro-Wilk and Kolmogorov-Smirnov tests to formally assess normality.
